---
title: "Gapminder Data Analysis"
author: ""
date: ""
output: html_document
---

```{r include=FALSE}
# global chunk options
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE, fig.align = "center")
```

## Load packages

```{r packages}
library(tidyverse)
```

## Load data

```{r data_read}
gap <- read_csv("https://bit.ly/gap_data")
```

## Exercises

### Exercise 1

Take a glimpse at the data set with function `glimpse()`.

```{r glimpse}
glimpse(gap)
```

How many variables and observations are in `gap`? What are the variable types
for the variables in `gap`? 6

Variable | Type
---------|----------
Var 1    |character
Var 2    |character
Var 3    |double
Var 4    |double
Var 5    |double
Var 6    |double

### Exercise 2

**Part 1**

Create a scatter plot of life expectancy versus year for the United States.

```{r ex2_part1, eval=FALSE}
gap_usa <- gap %>% 
  filter(country == "United States")

ggplot(data = gap_usa, mapping = aes(x = year, y = lifeExp)) +
  geom_point()
```

**Part 2**

Create a plot of life expectancy verses GDP per capita for Canada. Set the
color of the points to be red and connect them with a solid black line.

```{r ex2_part2}
# create plot here
gap_can <- gap %>%
   filter(country == "Canada")

ggplot(data = gap_can, mapping = aes(x = gdpPercap, y = lifeExp), line = solid) +
  geom_point(color = 'red')+
  geom_line()
```

**Part 3**

Create a scatter plot of life expectancy versus GDP per capita based on all 
countries in Africa in 1992. Set the color to be purple.

```{r ex2_part3}
# create plot here
gap_afr <- gap %>%
   filter(continent == "Africa"& year == 1992)

ggplot(data = gap_afr, mapping = aes(x = gdpPercap, y = lifeExp)) +
  geom_point(color = 'purple')
```

### Exercise 3

**Part 1**

Create a plot of life expectancy versus GDP per capita based on all the
countries in Africa and Europe in 1992. Set the point colors to code for the
two continents.

*Hints*:

1. Filter `gap` so it only contains observations from Africa or Europe.
   Do this with `continent %in% c("Africa", "Europe")`. Also, filter so that
   the observations are only from 1992. Separate multiple conditions with
   commas.
   
2. Add another aesthetic to the plot. Instead of just `x` and `y`, include
   `color = continent` inside function `aes()`. The color of the points will
   be set for each level of the variable `continent`.

```{r ex3_part1}
# create plot here
gap_ex3 <- gap %>%
   filter(continent == c("Africa","Europe")& year == 1992)

ggplot(data = gap_ex3, mapping = aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point()
```

**Part 2**

Create comparison boxplots by continent of GDP per capita for the year 2007.
Fill the boxplots with color `darkgreen`.

*Hints*:

1. Filter `gap` so it only contains observations from 2007.

2. Inside `aes()` set `x = continent, y = gdpPercap`.

3. Use `geom_boxplot()` and set the fill to `darkgreen`.

```{r ex3_part2}
# create plot here
gap_box <- gap %>%
   filter(year == 2007)

ggplot(data = gap_box, mapping = aes(x = continent, y = gdpPercap), line = solid) +
  geom_boxplot(fill = "darkgreen")
```
```{r ex3_part2}




```

